---
title: "R Notebook"
output: html_notebook
---

```{r include=FALSE}
library(sae)
library(tidycensus)
library(tidyverse)

census_api_key("9fb9cc8584a00b7006e7550c685129b65240b9aa")
```

```{r}
load("all_data.RDA")
all_data = all_data %>% select(-non_hispanic, -married, -other)

xi.large = all_data[, 8:ncol(all_data)] 
xi.small = xi.large %>% select(-asian, -male_only, -no_car)
```


```{r}
N = all_data %>% nrow()
gls = lm(povPerc~ white+ native + black  +  hispanic +
          female_only + degree + no_car + asian + male_only,
         data=all_data, weights = 1/(all_data$povPercSE^2))
gls %>% summary()
```


```{r}
xi = xi.large %>% as.matrix()
resultFH <- eblupFH(all_data$povPerc ~ xi, all_data$povPercSE^2, method="FH")
resultFH$fit$goodness

resultFH$fit$estcoef
```

```{r}
resultFH$fit$refvar
```


```{r}
data.frame(GLS = gls$fitted.values, y=all_data$povPerc, EB_FH=resultFH$eblup) %>% 
  gather(key="model", value="estimate", -y) %>% 
  ggplot(aes(x=log(y), y=log(estimate), color=model),  alpha=0.5) + geom_point() + 
  geom_abline(slope = 1, lty=2, alpha=0.5) + theme_bw()
```

```{r}
data.frame(gls = gls$fitted.values, y=all_data$povPerc, pop=all_data$pop,
           se = all_data$povPercSE, fh=resultFH$eblup) %>% 
        gather(key="model", value="estimate", -se, -y, -pop) %>% 
  ggplot(aes(x=-log(pop), y=y/estimate, color=model)) + geom_point() + 
  geom_hline(yintercept = 1, alpha=0.5, lty=2) + theme_bw()

```






```{r}
data.frame(gls = gls$fitted.values, y=all_data$povPerc, 
           se = all_data$povPercSE, fh=resultFH$eblup) %>% 
        gather(key="model", value="estimate", -se, -y) %>% 
  #filter(model=="fh") %>% 
  ggplot(aes(x=se, y=y/estimate, color=model)) + geom_point() + 
  geom_hline(yintercept = 1, alpha=0.5, lty=2) + theme_bw()

```

```{r}
data.frame(gls = gls$fitted.values, y=all_data$povPerc, 
           se = all_data$povPercSE, fh=resultFH$eblup) %>% 
        gather(key="model", value="estimate", -se, -y) %>% 
  filter(model=="fh") %>% 
  ggplot(aes(x=se, y=y/estimate, color=model)) + geom_point() + 
  geom_hline(yintercept = 1, alpha=0.5, lty=2) + theme_bw()

```



# Mapping 

Where do people live? 
```{r}
library(usmap)
df = all_data %>% select(fips, pop, povPerc, povPercSE) %>% 
  mutate(l_pop =log(pop), l_povPerc=log(povPerc), l_povPercSE = log(povPercSE))

plot_usmap("counties", values = "l_pop",
                  include = c("ND", "SD", "MN", "WI"), 
           data=df) + scale_fill_viridis_c() + theme(legend.position="bottom")
```

```{r}
df = all_data %>% select(fips, pop, povPerc, povPercSE) %>% 
  mutate(l_povPerc=log(povPerc), l_povPercSE = log(povPercSE), ratio=povPercSE/povPerc)

plot_usmap("counties", values = "l_povPerc",
                  include = c("ND", "SD", "MN", "WI"), 
           data=df) + scale_fill_viridis_c() + theme(legend.position="bottom")
```


```{r}
df = all_data %>% select(fips, pop, povPerc, povPercSE) %>% 
  mutate(l_povPerc=log(povPerc), l_povPercSE = log(povPercSE))

plot_usmap("counties", values = "povPercSE",
                  include = c("ND", "SD", "MN", "WI"), 
           data=df) + scale_fill_viridis_c() + theme(legend.position="bottom")
```

High SE/povPERC ratio

```{r}
df = all_data %>% select(fips, pop, povPerc, povPercSE) %>% 
  mutate(l_povPerc=log(povPerc), l_povPercSE = log(povPercSE), ratio=povPercSE/povPerc)

plot_usmap("counties", values = "ratio",
                  include = c("ND", "SD", "MN", "WI"), 
           data=df) + scale_fill_viridis_c() + theme(legend.position="bottom")
```



```{r}
test = data.frame(fips=all_data$fips, county=all_data$county,  gls = gls$fitted.values, y=all_data$povPerc, 
            se = all_data$povPercSE, fh=resultFH$eblup)%>% 
     gather(key="model", value="estimate", -se, -y, -fips, -county) %>% 
     filter(model=="fh") %>%  
    mutate(ratio = y/estimate) %>%
  select(fips, ratio)

plot_usmap(regions = "counties", values = "ratio", data=test,
                  include = c("ND", "SD")) + scale_fill_viridis_c() + theme(legend.position="bottom")
```

```{r}
df = all_data %>% select(fips, pop, povPerc, povPercSE) %>% 
  mutate(l_povPerc=log(povPerc), l_povPercSE = log(povPercSE))

plot_usmap("counties", values = "povPercSE",
                  include = c("ND", "SD"), 
           data=df) + scale_fill_viridis_c() + theme(legend.position="bottom")
```



# counties of interest (probably?)

TODO: FIND OUT WHERE THE DISAGREEMENTS WITH THE MODEL AND THE DIRECT ESTIMATES OCURR. 
Summarize differences between those places and the rest of the dataset. Is there a general trend?

```{r}
s = data.frame(fips=all_data$fips, county=all_data$county,  gls = gls$fitted.values, y=all_data$povPerc, 
            se = all_data$povPercSE, fh=resultFH$eblup)%>% 
     gather(key="model", value="estimate", -se, -y, -fips, -county) %>% 
     filter(model=="fh") %>% 
    mutate(ratio = y/estimate) %>% 
    filter( ratio > quantile(ratio, 0.99) | ratio < quantile(ratio, 0.01)  ) %>% arrange(desc(ratio))

s
```


46095  Mellette County, South Dakota
46117   Stanley County, South Dakota


46049     Faulk County, South Dakota
4 46071   Jackson County, South Dakota
5 46089 McPherson County, South Dakota


```{r}
df = all_data %>% 
  #filter(fips %in% c(46049, 46095, 46071, 46089)) 
  filter(fips %in% s$fips) 
df %>% select(-pov, -SE)
```

```{r}
s
```

```{r}
all_data %>% summary()
```

